---
title: "18-11-2024 course_part1"
author: "maulid_japhary"
date: "2024-11-18"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# course part 1 - Laboratory

we selected 20 isolates

- 6 **Klebsiella** spp. isolate

- 6 **Escherichia coli** isolates

- 8 **Salmonella** spp. isolates 
# course 2 Labworks (tuesday)
##Results of isolates species identification


![](`Files/VITECK RESULTS19_11_2024.pdf`)


picture of results
A1 E.coli
A4 E.coli
C4 Salmonella spp

Isolates were sub cultured for DNA Extraction were done on 19-11-2024
```{r}

```


```{r}

#install.packages("here")
#install.packages("DiagrammeR")
```

ctrl + alt + I (make a cell - shortcut on the keyboard)

```{r}
angry_or_not<-"not"
```
gives content of the object (container)
```{r}
angry_or_not
```




# working with your data


```{r}
here::here()
```


```{r}
path_data<-here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")
path_data
```
```{r}
human_data<-
  readr::read_csv2(path_data)
```







# Day 3
## packages into memory
```{r}
library(here)
library(tidyverse)
```



## Getting help - and what function do

F1


```{r}
?library()
```

## we get out data
```{r}
path_data<-here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")
path_data
```


```{r}
human_data <-
  readr::read_csv2(path_data)
```

## Exploration of a data frame 
- Names of columns
```{r}
colnames(human_data)
```

-getting the columns names into a vector for pasting and manipulations
- i can use a function within a function!
```{r}
dput(
  colnames(human_data)
)
```
-explanation:c() is to make vector
```{r}
test <-c("one","2",3)
test
test[3]
typeof(test[3])
typeof(test)

class(test)
```

```{r}
test2 <-c(1,2,3)
typeof(test2)
class(test2)
test2*test2
```
```{r}
test2*c(1,2)
```

-deactivate the output in the markdown document:in the document settings

near knit button:only select `chunk output in the console`

-renaming columns names (easier to use)

```{r}
original_names <- colnames(human_data)
original_names
```

```{r}
New_names <- str_replace_all(original_names, " ","_")
New_names
New_names[88]
```
 -regex pattern
```{r}
New_names2 <- str_replace_all(
  New_names, "[?/,;.*()-]","_")%>%
  str_remove_all("(^_*)|(_*$)")
New_names2
New_names2[88]
```
```{r}
colnames(human_data) <- New_names2
```

-selecting columns
```{r}
human_data_selection<-
human_data %>%
  
  select(INIKA_OH_TZ_ID, Age__yrs, Gender, Enter_a_date, Region, District, 
         Specify_if_other_district, Sample, Season, Origin_of_sample, 
         Which_class_grade_are_you, 
         Who_is_your_caretaker, 
         If_others__mention, 
         What_is_your_occupation_and_or_of_your_caretaker, 
         Have_you_ever_heard_about_AMR, If_yes__how_did_you_get_this_information, 
         Have_you_or_your_children_used_any_antibiotics_at_any_time, 
         If_yes__where_did_you_get_these_drugs_from,
         If_it_was_drug_sellers_or_pharmacy__did_you_have_a_prescription_from_the_doctor_prescriber,
         GPS_coordinates_latitude, GPS_coordinates_longitude) 

human_data_selection %>% View()

 head()
```
-> viewing your data or better of your data. Notion of pipe and selecting columns
```{r}
glimpse(human_data_selection)
str(human_data_selection)
```
#HERE
-something does not make sense - empty column
```{r}
human_data_selection <-
  human_data_selection %>%
  select(-Specify_if_other_district) %>%
  mutate_at(vars(Age__yrs) , as.integer) %>%
  mutate_if(is.character, factor) %>%
  mutate_at(vars(INIKA_OH_TZ_ID) , as.character) 
```


```{r}
levels(test$Region)
```
 -summary
```{r}
summary(human_data_selection)
```
 
```{r}
human_data_selection %>%
  filter(INIKA_OH_TZ_ID =="238")
```
```{r}
human_data_selection %>%
  filter(INIKA_OH_TZ_ID =="238") %>%
  filter(is.na(Gender))%>%
 view()
```
#summary()



