---
title: "18-11-2024 course_part1"
author: "maulid_japhary"
date: "2024-11-18"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# course part 1 - Laboratory

we selected 20 isolates

- 6 **Klebsiella** spp. isolate

- 6 **Escherichia coli** isolates

- 8 **Salmonella** spp. isolates 
# course 2 Labworks (tuesday)
##Results of isolates species identification


![](`Files/VITECK RESULTS19_11_2024.pdf`)


picture of results
A1 E.coli
A4 E.coli
C4 Salmonella spp

```{r}

```


```{r}

#install.packages("here")
#install.packages("DiagrammeR")
```

ctrl + alt + I (make a cell - shortcut on the keyboard)

```{r}
angry_or_not<-"not"
```
gives content of the object (container)
```{r}
angry_or_not
```




# working with your data


```{r}
here::here()
```


```{r}
path_data<-here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")
path_data
```
```{r}
human_data<-
  readr::read_csv2(path_data)
```







# Day 3
## packages into memory
```{r}
library(here)
library(tidyverse)
```



## Getting help - and what function do

F1


```{r}
?library()
```

## we get out data
```{r}
path_data<-here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")
path_data
```


```{r}
human_data <-
  readr::read_csv2(path_data)
```

## Exploration of a data frame 
- Names of columns
```{r}
colnames(human_data)
```

-getting the columns names into a vector for pasting and manipulations
- i can use a function within a function!
```{r}
dput(
  colnames(human_data)
)
```
-explanation:c() is to make vector
```{r}
test <-c("one","2",3)
test
test[3]
typeof(test[3])
typeof(test)

class(test)
```

```{r}
test2 <-c(1,2,3)
typeof(test2)
class(test2)
test2*test2
```
```{r}
test2*c(1,2)
```


